Modelling last glacial cycle ice dynamics in the Alps

### 2018-07-24 articles/cycle/alps-cycle.tex
Add maximum ice thickness and remove todo notes.

    \begin{figure*}[t]
      \centerline{\includegraphics{alpcyc_hr_inputs}}
      \caption{%
        \textbf{(a)} Geothermal heat flow from applying the similarity method
        to multiple geophysical proxies \citep{Goutorbe.etal.2011} used as a
        boundary condition to the bedrock thermal model 3\,km below the
        ice-bedrock interface.
        \textbf{(b)} Initial basal topography from SRTM
        \citep{Jarvis.etal.2008} and geomorphological reconstruction of Last
        Glacial Maximum Alpine glacier extent \citep[solid red
        line,][]{Ehlers.etal.2011}.
        \textbf{(c)} Extract from the estimated present-day ice thickness
        \citep{Huss.Farinotti.2012} substracted from the SRTM topography and
        aggregated to a 1\,km horizontal resolution.
        \textbf{(d)} Modern January and \textbf{(e)} July standard deviation
        \citep{Seguinot.2013} of daily mean temperature from the ERA-Interim
        \citep[1979--2012;][]{Dee.etal.2011} monthly climatology approximating
        temperature variability.
        \textbf{(f)} Modern January and \textbf{(g)} July mean near-surface air
        temperature, and \textbf{(h)} January and \textbf{(i)} July
        precipitation from WorldClim \citep[1960--1990;][]{Hijmans.etal.2005}
        used to compute surface mass balance. The background maps contain
        Natural Earth Data \citep{Patterson.Kelso.2017}.
        \todo{(R2) add scale and mark inset.}}
      \label{fig:inputs}
    \end{figure*}

@@

    \begin{figure*}[t]
      \centerline{\includegraphics{alpcyc_hr_lgmvel}}
      \caption{%
        \textbf{(a)} Modelled bedrock topography (grey), ice surface topography
        (200\,m contours), and ice surface velocity (blue) in the Alps
        24.57\,ka before present, corresponding to the maximum modelled ice
        cover. The modelled LGM ice extent (dashed orange line) and
        geomorphological reconstruction \citep[solid red
        line,][]{Ehlers.etal.2011}, and some major transfluences across
        mountain passes (crosses) are shown.
        \textbf{(b)} Temperature offset time-series from the EPICA ice core
        used as palaeo-climate forcing for the ice flow model \citep[black
        curve,][]{Jouzel.etal.2007}, and modelled total ice volume through the
        last glacial cycle (120--0\,ka), expressed in meters of sea level
        equivalent (m~s.l.e., blue curve). Shaded gray areas indicate the
        timing for Marine Oxygen Isotope Stage (MIS)~2 and MIS~4 according to
        a~global compilation of benthic \chem{\delta^{18}O} records
        \citep{Lisiecki.Raymo.2005}.
        \todo{Thicken the 200\,m contours on all figures. Remove ice domes and
              replace Straniger.}}
      \label{fig:lgmvel}
    \end{figure*}

@@

    \begin{figure*}[t]
      \centerline{\includegraphics{alpcyc_hr_timing}}
      \caption{%
        \textbf{(a)} Timing of the LGM given by the modelled age of maximum ice
        thickness throughout the entire simulation (colour mapping) and
        corresponding, ice surface elevation (200\,m contours).
        \textbf{(b)} Temperature offset time-series from the EPICA ice core
        used as palaeo-climate forcing for the ice flow model (black curve),
        and modelled glacierized area during the LGM (coloured curve). The LGM
        is here modelled as a time-transgressive event. In much of the mountain
        area, maximum thickness is reached before 27\,ka.
        \todo{fix colourbar label}}
      \label{fig:timing}
    \end{figure*}

@@

    In the upper Rhone Valley, the maximum ice surface elevation reached
    during MIS~2 in our simulation, corrected for bedrock deformation, is
    consistently modelled to have occured several hundred metres above the
    observed trimline elevations (Fig.~\ref{fig:trimlines}a and b), with a mean
    difference of $861\pm197$\,m (one standard deviation,
    Fig.~\ref{fig:trimlines}b), wich is significant in respect to the modelled
    regional ice thickness \todo{add a number}. The result is somewhat
    dependent on uncertain basal sliding and ice rheological parameters
    to which the model sensitivity was not tested here. However, a previous
    sensitivity study conducted with PISM and a very similar model set-up on
    the Cordilleran ice sheet show that varying ice rheological parameters
    resulted in relative errors in modelled total ice volume of 25\,\unit{\%}
    with regard to ice rheological parameters and 14\,\unit{\%} regarding basal
    sliding parameters \citep[Fig.~7]{Seguinot.etal.2016}. In addition,
    regional sensitivity tests in the Rhone Valley show that a modelled ice
    surface compatible with the trimline elevations is incompatible with the
    documented overspilling flow to the south across the Simplon pass
    \citep{Becker.etal.2017}.

>>>>

    \begin{figure*}[t]
      \centerline{\includegraphics{alpcyc_hr_inputs}}
      \caption{%
        \textbf{(a)} Geothermal heat flow from applying the similarity method
        to multiple geophysical proxies \citep{Goutorbe.etal.2011} used as a
        boundary condition to the bedrock thermal model 3\,km below the
        ice-bedrock interface.
        \textbf{(b)} Initial basal topography from SRTM
        \citep{Jarvis.etal.2008} and geomorphological reconstruction of Last
        Glacial Maximum Alpine glacier extent \citep[solid red
        line,][]{Ehlers.etal.2011}.
        \textbf{(c)} Extract from the estimated present-day ice thickness
        \citep{Huss.Farinotti.2012} substracted from the SRTM topography and
        aggregated to a 1\,km horizontal resolution.
        \textbf{(d)} Modern January and \textbf{(e)} July standard deviation
        \citep{Seguinot.2013} of daily mean temperature from the ERA-Interim
        \citep[1979--2012;][]{Dee.etal.2011} monthly climatology approximating
        temperature variability.
        \textbf{(f)} Modern January and \textbf{(g)} July mean near-surface air
        temperature, and \textbf{(h)} January and \textbf{(i)} July
        precipitation from WorldClim \citep[1960--1990;][]{Hijmans.etal.2005}
        used to compute surface mass balance. The background maps contain
        Natural Earth Data \citep{Patterson.Kelso.2017}.}
      \label{fig:inputs}
    \end{figure*}

@@

    \begin{figure*}[t]
      \centerline{\includegraphics{alpcyc_hr_lgmvel}}
      \caption{%
        \textbf{(a)} Modelled bedrock topography (grey), ice surface topography
        (200\,m contours), and ice surface velocity (blue) in the Alps
        24.57\,ka before present, corresponding to the maximum modelled ice
        cover. The modelled LGM ice extent (dashed orange line) and
        geomorphological reconstruction \citep[solid red
        line,][]{Ehlers.etal.2011}, and some major transfluences across
        mountain passes (crosses) are shown.
        \textbf{(b)} Temperature offset time-series from the EPICA ice core
        used as palaeo-climate forcing for the ice flow model \citep[black
        curve,][]{Jouzel.etal.2007}, and modelled total ice volume through the
        last glacial cycle (120--0\,ka), expressed in meters of sea level
        equivalent (m~s.l.e., blue curve). Shaded gray areas indicate the
        timing for Marine Oxygen Isotope Stage (MIS)~2 and MIS~4 according to
        a~global compilation of benthic \chem{\delta^{18}O} records
        \citep{Lisiecki.Raymo.2005}.}
      \label{fig:lgmvel}
    \end{figure*}

@@

    \begin{figure*}[t]
      \centerline{\includegraphics{alpcyc_hr_timing}}
      \caption{%
        \textbf{(a)} Timing of the LGM given by the modelled age of maximum ice
        thickness throughout the entire simulation (colour mapping) and
        corresponding, ice surface elevation (200\,m contours).
        \textbf{(b)} Temperature offset time-series from the EPICA ice core
        used as palaeo-climate forcing for the ice flow model (black curve),
        and modelled glacierized area during the LGM (coloured curve). The LGM
        is here modelled as a time-transgressive event. In much of the mountain
        area, maximum thickness is reached before 27\,ka.}
      \label{fig:timing}
    \end{figure*}

@@

    In the upper Rhone Valley, the maximum ice surface elevation reached
    during MIS~2 in our simulation, corrected for bedrock deformation, is
    consistently modelled to have occured several hundred metres above the
    observed trimline elevations (Fig.~\ref{fig:trimlines}a and b), with a mean
    difference of $861\pm197$\,m (one standard deviation,
    Fig.~\ref{fig:trimlines}b), which is significant in respect to the modelled
    maximum ice thickness in the Rhone Valley (and the Alps) of 2586\,m. This
    result depends on uncertain basal sliding and ice rheological parameters
    to which the model sensitivity was not tested here. However, a previous
    sensitivity study conducted with PISM and a very similar model set-up on
    the Cordilleran ice sheet show that varying ice rheological parameters
    resulted in relative errors in modelled total ice volume of 25\,\unit{\%}
    with regard to ice rheological parameters and 14\,\unit{\%} regarding basal
    sliding parameters \citep[Fig.~7]{Seguinot.etal.2016}. In addition,
    regional sensitivity tests in the Rhone Valley show that a modelled ice
    surface compatible with the trimline elevations is incompatible with the
    documented overspilling flow to the south across the Simplon pass
    \citep{Becker.etal.2017}.
